---
title: "Soy Endophytes CO2 Full Code"
author: "Connor Morozumi"
date: "most recent edits `r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
    code_folding: hide
    theme: cosmo
    df_print: paged
---

# Load pkgs
```{r}
library(tidyverse)
library(here)
library(patchwork)
library(cowplot)
library(ggVennDiagram)
library(ggpubr)

library(vegan)
library(phyloseq)
library(SpiecEasi)
library(NetCoMi)
```


# Read in data
```{r}
cleaned.files= here::here("datasets")
net.loc= here("network_analysis")
output.loc= here("outputs")

#its1
its1=read.csv(file= paste0(cleaned.files,"/cleaned.ASVs.ITS1.csv"), row.names = "X")

its1.net= read_csv(file= paste0(cleaned.files,"/cleaned.ASVs.ITS1.csv"))
#its2
its2 = read.csv(file= paste0(cleaned.files, "/cleaned.ASVs.its2.csv"), row.names = "X")

its2.net = read_csv(file= paste0(cleaned.files, "/cleaned.ASVs.its2.csv"))
#bacteria
bac= read.csv(file= paste0(cleaned.files,"/cleaned.ASVs.bacteria.csv"), row.names = "X")

bac.net= read_csv(file= paste0(cleaned.files,"/cleaned.ASVs.bacteria.csv"))

# two additional ones for higher level taxa summaries
# bac.asvnos= read.csv(file= paste0(cleaned.files,"/cleaned.nothost.bacteria.asv_labeled_by_asv_no.csv"), row.names = "X")
# 
# tax dfs
bac.tax= read.csv(file= paste0(cleaned.files,"/cleaned.nothost.bacteria.taxa.matched_to_asv_no.csv"), row.names = "X")

its1.tax= read.csv(file= paste0(cleaned.files,"/cleaned.nothost.its1.taxa.matched_to_asv_no.csv"), row.names = "X")

its2.tax= read.csv(file= paste0(cleaned.files,"/cleaned.nothost.its2.taxa.matched_to_asv_no.csv"), row.names = "X")

#metadata
meta= read.csv(file= paste0(cleaned.files, "/cleaned.meta.csv"), sep=",", header = T)

```

## Cleaning
```{r}

# change plot names for congruence with metadata and taxa dfs
meta$site= sub("_", "-", meta$site)

its1= its1 %>% 
  rownames_to_column("site")

its1$site= sub("_", "-", its1$site)

its1 = its1 %>% 
  column_to_rownames("site")

meta.lite= meta %>% 
  dplyr::select(site, CO2, plot, rep)

# Clean up family names for more readable legend
its1.tax$Family.legend= gsub(".*__","",its1.tax$Family)

its2.tax$Family.legend= gsub(".*__","",its2.tax$Family)

# sample ID
# fix sample names
names(its1.net)[1] ="Sample"

its1.net$Sample= sub("_", "-", its1.net$Sample)

names(its2.net)[1] ="Sample"

names(bac.net)[1] ="Sample"
```

# SUMMARY STATS

## Read counts
```{r}
# link meta to asv dfs too

## its1
read.counts.its1= its1 %>% 
  rownames_to_column("site") %>% 
  full_join(., meta.lite) %>% 
  dplyr::select(-c(plot, rep)) %>% 
  group_split(CO2)

names(read.counts.its1)<- c("control", "elevated")

read.counts.its1= map(read.counts.its1, dplyr::select,!CO2)

read.counts.its1= map(read.counts.its1, column_to_rownames, "site")

# filter out columns that are completely blank, these can be induced once the dataset is split by treatment
no.zeros.its1=list()
no.zeros.its1$control=read.counts.its1$control[colSums(read.counts.its1$control, na.rm = TRUE) > 0]
no.zeros.its1$elevated=read.counts.its1$elevated[colSums(read.counts.its1$elevated, na.rm = TRUE) > 0]

## bac
read.counts.bac= bac %>% 
  rownames_to_column("site") %>% 
  full_join(., meta.lite) %>% 
  dplyr::select(-c( plot, rep)) %>% 
  group_split(CO2)

names(read.counts.bac)<- c("control", "elevated")

read.counts.bac= map(read.counts.bac, dplyr::select,!CO2)

read.counts.bac= map(read.counts.bac, column_to_rownames, "site")


no.zeros.bac=list()
no.zeros.bac$control=read.counts.bac$control[colSums(read.counts.bac$control, na.rm = TRUE) > 0]
no.zeros.bac$elevated=read.counts.bac$elevated[colSums(read.counts.bac$elevated, na.rm = TRUE) > 0]

## its2
read.counts.its2= its2 %>% 
  rownames_to_column("site") %>% 
  full_join(., meta.lite) %>% 
  dplyr::select(-c(plot, rep)) %>% 
  group_split(CO2)

names(read.counts.its2)<- c("control", "elevated")

read.counts.its2= map(read.counts.its2, dplyr::select,!CO2)

read.counts.its2= map(read.counts.its2, column_to_rownames, "site")

map(read.counts.its2, sum)

no.zeros.its2=list()
no.zeros.its2$control=read.counts.its2$control[colSums(read.counts.its2$control, na.rm = TRUE) > 0]
no.zeros.its2$elevated=read.counts.its2$elevated[colSums(read.counts.its2$elevated, na.rm = TRUE) > 0]
```

```{r}
## sum read counts across primers 

its1.read.sums= map(no.zeros.its1, sum)
its2.read.sums= map(no.zeros.its2, sum)
bac.read.sums= map(no.zeros.bac, sum)

its1.read.sums.tot= sum(unlist(its1.read.sums, recursive=FALSE))
its2.read.sums.tot= sum(unlist(its2.read.sums, recursive=FALSE))
bac.read.sums.tot= sum(unlist(bac.read.sums, recursive=FALSE))
```

# COMMUNITY METRICS
## Richness and diversity

Per sample
```{r}
#richness per sample
bac$richness= rowSums(bac > 0)

its1$richness= rowSums(its1 > 0)

its2$richness= rowSums(its2 > 0)

# shannon, simp

bac$H = diversity(bac)
its1$H = diversity(its1)
its2$H = diversity(its2)

bac$simp = diversity(bac, "simpson")
its1$simp = diversity(its1, "simpson")
its2$simp = diversity(its2, "simpson")
```

Per plot
To calculate per plot richness we are going to have to do some extra massaging
```{r}
plot.rich.bac= bac %>% 
  rownames_to_column("site")
plot.rich.its1= its1 %>% 
  rownames_to_column("site")
plot.rich.its2= its2 %>% 
  rownames_to_column("site")

plot.rich.bac=full_join(plot.rich.bac, meta.lite)
plot.rich.its1=full_join(plot.rich.its1, meta.lite)
plot.rich.its2=full_join(plot.rich.its2, meta.lite)

```

Link to metadata
```{r}
# rownames to cols to link meta data in
bac.div= bac %>% 
  dplyr::select(richness, H, simp) %>% 
  rownames_to_column("site")

its1.div= its1 %>% 
  dplyr::select(richness, H, simp) %>% 
  rownames_to_column("site")

its2.div= its2 %>% 
  dplyr::select(richness, H, simp) %>% 
  rownames_to_column("site")


# link to meta
bac.div= full_join(bac.div, meta.lite)

its1.div= full_join(its1.div, meta.lite)

its2.div= full_join(its2.div, meta.lite)
```


## Composition

Transform data to proportions or apply Hellinger

```{r}
ITS1.asv.prop=prop.table(as.matrix(its1),1)

ITS1.asv.hell <- decostand(its1, "hellinger")

its2.asv.prop=prop.table(as.matrix(its2),1)

its2.asv.hell <- decostand(its2, "hellinger")

bacteria.asv.prop=prop.table(as.matrix(bac),1)

bacteria.asv.hell <- decostand(bac, "hellinger")

```


Visualization

NMDS for each primer on the transformed mat

ITS1
```{r}
its1.mds <- metaMDS(ITS1.asv.prop)
its1.mds.mds.3 <- metaMDS(ITS1.asv.prop, k=3)
metaMDS(ITS1.asv.prop, k=3, previous.best = its1.mds.mds.3)

#hellinger
its1.mds.hell <- metaMDS(ITS1.asv.hell)
its1.mds.3.hell <- metaMDS(ITS1.asv.hell, k=3)
metaMDS(ITS1.asv.hell, k=3, previous.best = its1.mds.3.hell)
```
A three dim `k=3` solution is better

```{r}
nmds.its1= its1.mds.mds.3 ## CHANGE ME
#site scores
data.scores.its1 <- as.data.frame(scores(nmds.its1)$sites)  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores.its1$site <- rownames(data.scores.its1)  # create a column of site names, from the rownames of data.scores
data.scores.its1=full_join(data.scores.its1, meta)

#species scores
species.scores.its1 <- as.data.frame(scores(nmds.its1, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores.its1$species <- rownames(species.scores.its1)  # create a column of species, from the rownames of species.scores
```

```{r}
its1.A= ggplot(data= data.scores.its1,aes(x=NMDS1,y=NMDS2,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.its1,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.its1,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)

if("NMDS3" %in% colnames(data.scores.its1) ==TRUE){
its1.B=ggplot(data= data.scores.its1,aes(x=NMDS1,y=NMDS3,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.its1,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.its1,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)
}

#axis 2 vs 3
if("NMDS3" %in% colnames(data.scores.its1) ==TRUE){
its1.C=ggplot(data= data.scores.its1,aes(x=NMDS2,y=NMDS3,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.its1,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.its1,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)
}

its1.all= its1.A + its1.B + its1.C
```


its2

```{r}

its2.mds.hell <- metaMDS(its2.asv.hell)
its2.mds.hell.3 <- metaMDS(its2.asv.hell, k=3)
metaMDS(its2.asv.hell, k=3, previous.best = its2.mds.hell.3)


its2.mds.prop <- metaMDS(its2.asv.prop)
its2.mds.prop.3 <- metaMDS(its2.asv.prop, k=3)
nmds.its2=metaMDS(its2.asv.prop, k=3, previous.best = its2.mds.prop.3)

```

```{r}

#site scores
data.scores.its2 <- as.data.frame(scores(nmds.its2)$sites)  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores.its2$site <- rownames(data.scores.its2)  # create a column of site names, from the rownames of data.scores

data.scores.its2=full_join(data.scores.its2, meta)

#species scores
species.scores.its2 <- as.data.frame(scores(nmds.its2, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores.its2$species <- rownames(species.scores.its2)  # create a column of species, from the rownames of species.scores
```

```{r}
its2.A=ggplot(data= data.scores.its2,aes(x=NMDS1,y=NMDS2,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.its2,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.its2,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)

#axis 1 v 3

if("NMDS3" %in% colnames(data.scores.its2) ==TRUE){
its2.B=ggplot(data= data.scores.its2,aes(x=NMDS1,y=NMDS3,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.its2,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.its2,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)
}

#axis 2 vs 3
if("NMDS3" %in% colnames(data.scores.its2) ==TRUE){
its2.C=ggplot(data= data.scores.its2,aes(x=NMDS2,y=NMDS3,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.its2,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.its2,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)
}

its2.all= its2.A + its2.B + its2.C

```

Bacteria
```{r}

# NMDS

bacteria.mds.hell <- metaMDS(bacteria.asv.hell)
bacteria.mds.hell.3 <- metaMDS(bacteria.asv.hell, k=3)
metaMDS(bacteria.asv.hell, k=3, previous.best = bacteria.mds.hell.3)


bacteria.mds.prop <- metaMDS(bacteria.asv.prop)
bacteria.mds.prop.3 <- metaMDS(bacteria.asv.prop, k=3)
nmds.bac=metaMDS(bacteria.asv.hell, k=3, previous.best = bacteria.mds.hell.3)
  # no convergence k=2
```

```{r}

#site scores
data.scores.bac <- as.data.frame(scores(nmds.bac)$sites)  #Using the scores function from vegan to extract the site scores and convert to a data.frame
data.scores.bac$site <- rownames(data.scores.bac)  # create a column of site names, from the rownames of data.scores

data.scores.bac=full_join(data.scores.bac, meta)

#species scores
species.scores.bac <- as.data.frame(scores(nmds.bac, "species"))  #Using the scores function from vegan to extract the species scores and convert to a data.frame
species.scores.bac$species <- rownames(species.scores.bac)  # create a column of species, from the rownames of species.scores
```

```{r}
bac.a=ggplot(data= data.scores.bac,aes(x=NMDS1,y=NMDS2,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.bac,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.bac,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue"), labels= c("Elevated", "Ambient")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)

#axis 1 v 3

if("NMDS3" %in% colnames(data.scores.bac) ==TRUE){
bac.b=ggplot(data= data.scores.bac,aes(x=NMDS1,y=NMDS3,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.bac,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.bac,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)
}

#axis 2 vs 3
if("NMDS3" %in% colnames(data.scores.bac) ==TRUE){
bac.c= ggplot(data= data.scores.bac,aes(x=NMDS2,y=NMDS3,shape= CO2,colour=CO2)) + 
  #geom_text(data=species.scores.bac,aes(x=NMDS1,y=NMDS2,label=species),alpha=0.5) +  # add the species labels
  geom_point(size=3) + # add the point markers
  #geom_text(data=data.scores.bac,aes(x=NMDS1,y=NMDS2,label=site),size=6,vjust=0) +  # add the site labels
  scale_colour_manual(values=c("Elevated" = "red", "Control" = "blue")) +
  coord_equal() +
  theme_bw()+
  guides(color=guide_legend(expression(paste("CO"["2"]))), shape = FALSE)
}

bac.all= bac.a + bac.b + bac.c

```

Fig 1: multiplot fungi and bacterial

```{r fig.width= 14}
its2.A.noleg= its2.A + theme(legend.position = "none") + geom_label(aes(x = 0.3, y = 1.5, label = "Fungal composition"), inherit.aes = F)

bac.a.bot= bac.a + theme(legend.position = "bottom") + geom_label(aes(x = -.2, y = 1, label = "Bacterial composition"), inherit.aes = F) + expand_limits(x = 0.7, y= 1.2)

fig1.vertical= its2.A.noleg / bac.a.bot 

ggsave(paste0(output.loc, "/fig.1.pdf"), fig1.vertical, width = 6)
ggsave(paste0(output.loc, "/fig.1.jpg"), fig1.vertical, width = 6)

```

### Stats

ITS1
```{r}
# plot 20_9 doesn't appear to be in the ASV tab and is blank in the meta tab
meta.fin=meta %>% 
  filter(site !="20_9")
adonmod=adonis2(ITS1.asv.hell ~ CO2, data = meta.fin)


# check for dispersion
  #use data.scores df to get corresponding treatment assignments in proper order
treatment= data.scores.its1 %>% 
  filter(site!= "20_9")

dis.its1= vegdist(ITS1.asv.hell)
mod <- betadisper(dis.its1, treatment$CO2)
anova(mod)
(mod.HSD <- TukeyHSD(mod))
plot(mod.HSD)
```

its2
```{r}
adonmod_ITS2=adonis2(its2.asv.prop ~ CO2, data = meta.fin)

# check for dispersion
  #use data.scores df to get corresponding treatment assignments in proper order
treatment= data.scores.its2 %>% 
  filter(site!= "20_9")

dis= vegdist(its2.asv.prop)
mod <- betadisper(dis, treatment$CO2)
anova(mod)
(mod.HSD <- TukeyHSD(mod))
plot(mod.HSD)
```

Bacteria

```{r}

# adonis
adonmod_16s=adonis2(bacteria.asv.hell ~ CO2, data = meta.fin)

# check for dispersion
  #use data.scores df to get corresponding treatment assignments in proper order
treatment= data.scores.bac %>% 
  filter(site!= "20-9")

dis= vegdist(bacteria.asv.hell)
mod <- betadisper(dis, treatment$CO2)
anova(mod)
(mod.HSD <- TukeyHSD(mod))
plot(mod.HSD)
```

# NETWORKS

- Filtering low abundant ASVs, maintaining ASVs with more than 10 read counts

Remove low abundant ASVs
```{r}

# ITS1
 its1.net= its1.net %>% column_to_rownames("Sample")
its1.sub=its1.net[ ,colSums(its1.net[,1:ncol(its1.net)]) >= 10]
paste("filtering low abundant ASVs removes", ncol(its1.net)- ncol(its1.sub), "ASVs")

# its2
its2.net= its2.net %>% column_to_rownames("Sample")
its2.sub=its2.net[ ,colSums(its2.net[,1:ncol(its2.net)]) >= 10]
paste("filtering low abundant ASVs removes", ncol(its2.net)- ncol(its2.sub), "ASVs")

# 16s
bac.net= bac.net %>% column_to_rownames("Sample")
bac.sub=bac.net[ ,colSums(bac.net[,2:ncol(bac.net)]) >= 10]
paste("filtering low abundant ASVs removes", ncol(bac.net)- ncol(bac.sub), "ASVs")

```

Filter all.tax by the filtered asv tables
```{r}
its1.sub.vec= names(its1.sub)

its1.tax.sub= its1.tax %>% 
  filter(uniqueID %in% its1.sub.vec)

its2.sub.vec= names(its2.sub)

its2.tax.sub= its2.tax %>% 
  filter(uniqueID %in% its2.sub.vec)

bac.sub.vec= names(bac.sub)

bac.tax.sub= bac.tax %>% 
  filter(uniqueID %in% bac.sub.vec)
```

Split Elevated CO~2~ and Ambient

Combine 3 asv dfs into a list item
```{r}
# all.asvs= list(ITS1.asv.filter, its2.asv.filter, bacteria.asv.tab.filter) #old
all.asvs= list(its1.sub, its2.sub, bac.sub)
names(all.asvs)<- c("its1", "its2", "16s")

# all.tax= list(its1.tax, its2.tax, bac.tax) #old
all.tax= list(its1.tax.sub, its2.tax.sub, bac.tax.sub)

names(all.tax)<- c("its1", "its2", "16s")
```

```{r}
# make a col named sample to link metadata in
all.asvs=all.asvs %>% 
map(rownames_to_column, "Sample")

# add metadata 
  
meta$site= sub("_", "-", meta$site)

meta.lite= meta %>% 
  select(site, CO2) %>% 
  rename(Sample= site)

all.asvs= all.asvs %>% 
  map(full_join, meta.lite) %>% # add metadata to asv data so we know which Samples are Elevated and which are Control
  map(relocate, "CO2", .after= "Sample") # move co2 treatment to start of df for each so we can see it

Elevated=all.asvs %>% 
  map(filter, CO2 == "Elevated") %>% 
  map(select, -c(CO2)) %>% 
  map( column_to_rownames, "Sample")

# Elevated[1] # check to see if that worked
Elevated[3]

Ambient=all.asvs %>% 
  map(filter, CO2 == "Control") %>% 
  map(select, -c(CO2)) %>% 
  map( column_to_rownames, "Sample")

# dumb way to do this, def better way to go about it. Stitch back together as a list
control.amb= c(Elevated, Ambient)
names(control.amb)<- c("elevated.its1", "elevated.its2", "elevated.16s", "amb.its1", "amb.its2", "amb.16s")
```

Make phyloseq objects of these networks
```{r}
# make phyloseq otu tables 
otu= map(control.amb, otu_table, taxa_are_rows = F)

all.tax= all.tax %>% 
  map(.,~rownames_to_column(.x, var="seq"))

all.tax= map(all.tax, column_to_rownames,var= "uniqueID")

all.tax=map(all.tax, select, -"seq")


# duplicated these into elevated and ambient lists
Ambient.tax=all.tax 
Elevated.tax=all.tax 

# dumb way to do this, def better way to go about it. Stitch back together as a list
control.amb.tax= c(Elevated.tax, Ambient.tax)
names(control.amb.tax)<- c("elevated.its1", "elevated.its2", "elevated.16s", "amb.its1", "amb.its2", "amb.16s")


## make tax object

control.amb.tax= map(control.amb.tax, as.matrix)

TAX = map(control.amb.tax, tax_table)

# combine into phyloseq obj
phylo.all= map2(otu, TAX, phyloseq)
```

## Create networks via SpeicEasi

From phyloseq obj
```{r}
mapply(assign, names(phylo.all), phylo.all, MoreArgs=list(envir = globalenv()))
```

```{r}
#its1
# gen.se.elevated.its1 <- spiec.easi(elevated.its1, method='mb', lambda.min.ratio=1e-2, nlambda=100, pulsar.params=list(rep.num=50))
# 
# ## check stability 
# getStability(gen.se.elevated.its1) #0.045 that seems good enough
# 
# # 
# gen.se.amb.its1 <- spiec.easi(amb.its1, method='mb', lambda.min.ratio=1e-2, nlambda=100, pulsar.params=list(rep.num=50))
# 
# ## check stability 
# getStability(gen.se.amb.its1) #0.0426
# 
# #its2 
# pargs2 <- list(rep.num=50, seed=10010, ncores=4)
# 
# se.elevated.its2.bstars <- spiec.easi(elevated.its2, method='mb', lambda.min.ratio=1e-1, nlambda=100,
#                sel.criterion='bstars', pulsar.select=TRUE, pulsar.params=pargs2)
# 
# getStability(se.elevated.its2.bstars)  # 
# 
# save(se.elevated.its2.bstars, file = paste0(net.loc, "/spiec_nets_its2_ele.bstars.RData"))

# se.amb.its2.bstars <- spiec.easi(elevated.its2, method='mb', lambda.min.ratio=1e-1, nlambda=100,
#                sel.criterion='bstars', pulsar.select=TRUE, pulsar.params=pargs2)
# 
# getStability(se.amb.its2.bstars)  # 
# 


# bac
# gen.se.elevated.bac <- spiec.easi(elevated.16s, method='mb', lambda.min.ratio=1e-2, nlambda=100, pulsar.params=list(rep.num=50))

# changing lambda min ration because stability is kinda low. Alternative could be to raise nlambda but 100 was what the troublehooting guide recommended on the high side

# gen.se.elevated.bac <- spiec.easi(elevated.16s, method='mb', lambda.min.ratio=1e-1, nlambda=100, pulsar.params=list(rep.num=50))
# 
# getStability(gen.se.elevated.bac) #  0.0462359 better!
# gen.se.elevated.bac$select$stars$summary
# 
# # 
# gen.se.amb.bac <- spiec.easi(amb.16s, method='mb', lambda.min.ratio=1e-1, nlambda=100, pulsar.params=list(rep.num=50))
# 
# getStability(gen.se.amb.bac) # 0.0475

```

## Save networks / work from a saved copy if necessary
```{r}
#save(se.elevated.bac, se.amb.bac, file = paste0(net.loc, "/spiec_nets_16s.RData"))

#save(se.elevated.its1, se.amb.its1, file = paste0(net.loc, "spiec_nets_ITS1.RData"))

#save(se.amb.its2.bstars, se.elevated.its2.bstars, file = paste0(net.loc, "/spiec_nets_its2.RData"))


if(!"se.elevated.its1" %in% names(globalenv()))
{load(paste0(net.loc,"/spiec_nets_ITS1.RData"))}

if(!"se.elevated.bac" %in% names(globalenv()))
{load(paste0(net.loc,"/spiec_nets_16s.RData"))}

if(!"se.elevated.its2.bstars" %in% names(globalenv()))
{load(paste0(net.loc,"/spiec_nets_its2.RData"))}

```


## Plot networks

ITS1

Elevated 
```{r}
# plot within phyloseq framework

its1.ele.mb <- adj2igraph(getRefit(se.elevated.its1),  vertex.attr=list(name=taxa_names(phylo.all$elevated.its1)))

p.its1.el= plot_network(its1.ele.mb, phylo.all$elevated.its1, type='taxa', color="Family.legend", label= NULL, point_size = 0.5, alpha= 0.5, line_alpha = 0.4, title = expression(paste("Elevated ", CO[2])))

# can do further ggplot functions 
p.its1.el= p.its1.el + 
  theme(legend.position = "none")
```

Ambient
```{r}
its1.amb.mb <- adj2igraph(getRefit(se.amb.its1),  vertex.attr=list(name=taxa_names(phylo.all$amb.its1)))

p.its1.amb= plot_network(its1.amb.mb, phylo.all$amb.its1, type='taxa', color="Family.legend", label= NULL, point_size = 0.5, alpha= 0.5, line_alpha = 0.4, title = expression(paste("Ambient ", CO[2])))

p.its1.amb.2= p.its1.amb + 
  guides(color=guide_legend(ncol=2)) +
  theme(legend.position = "none") 

test= p.its1.amb + 
 labs(color = "") +
  theme_void(base_size = 10) + guides(color = guide_legend(override.aes = list(size=3)))
```

```{r fig.width=15, fig.height=10}
# Extract the legend. Returns a gtable
leg <- ggpubr::get_legend(test)

# Convert to a ggplot and print
leg.plot=as_ggplot(leg) 
```

Make a multiplot
```{r}
p1= plot_grid(p.its1.amb.2, p.its1.el, labels = c('A', 'B'), hjust = 0.5)

p1.final= plot_grid(p1, NULL, leg.plot, ncol=1, align="v", rel_heights = c(1, 0, 1))

ggsave(paste0(output.loc,"/figureS3.jpg"),p1.final, height= 10)
ggsave(paste0(output.loc,"/figureS3.pdf"),p1.final, height= 10)
```

ITS2

```{r}
# plot within phyloseq framework

el.its2.mb <- adj2igraph(getRefit(se.elevated.its2.bstars),  vertex.attr=list(name=taxa_names(phylo.all$elevated.its2)))

p.its2.el= plot_network(el.its2.mb, phylo.all$elevated.its2, type='taxa', color="Family.legend", label= NULL, point_size = 0.5, alpha= 0.5, line_alpha = 0.4, title = expression(paste("Elevated ", CO[2])))

# can do further ggplot functions 
p.its2.el= p.its2.el + 
  theme(legend.position = "none")
```

```{r}
am.its2.mb <- adj2igraph(getRefit(se.amb.its2.bstars),  vertex.attr=list(name=taxa_names(phylo.all$amb.its2)))

p.its2.am= plot_network(am.its2.mb, phylo.all$amb.its2, type='taxa', color="Family.legend", label= NULL, point_size = 0.5, alpha= 0.5, line_alpha = 0.4, title = expression(paste("Ambient ", CO[2])))

# can do further ggplot functions 
p.its2.amb.2= p.its2.am + 
  guides(color=guide_legend(ncol=2)) +
  theme(legend.position = "none") 

#for the legend
leg.its2= p.its2.am + 
 labs(color = "") +
  theme_void(base_size = 10) + guides(color = guide_legend(override.aes = list(size=3)))
```

```{r fig.width=15, fig.height=10}
# Extract the legend. Returns a gtable
leg.its2 <- ggpubr::get_legend(leg.its2)

# Convert to a ggplot and print
leg.its2.plot=ggpubr::as_ggplot(leg.its2) 
```

Make a multiplot
```{r}
p1.its2= plot_grid(p.its2.amb.2, p.its2.el, labels = c('A', 'B'), hjust = 0.5)

p1.its2.final= plot_grid(p1.its2, NULL, leg.its2.plot, ncol=1, align="v", rel_heights = c(1, 0, 1))

ggsave(paste0(output.loc,"/figure2.jpg"),p1.its2.final, height= 10)
ggsave(paste0(output.loc,"/figure2.pdf"),p1.its2.final, height= 10)

```

16s
```{r}
#bac
bac.el.mb <- adj2igraph(getRefit(se.elevated.bac),  vertex.attr=list(name=taxa_names(phylo.all$elevated.16s)))

p.16s.el= plot_network(bac.el.mb, phylo.all$elevated.16s, type='taxa', color="Genus", point_size = 1, alpha= 0.5, line_alpha = 0.4, label= NULL, title = expression(paste("Elevated ", CO[2])))

p.16s.el= p.16s.el + theme(legend.position = "none")

```


``` {r}
bac.amb.mb <- adj2igraph(getRefit(se.amb.bac),  vertex.attr=list(name=taxa_names(phylo.all$amb.16s)))

p.16s.amb= plot_network(bac.amb.mb, phylo.all$amb.16s, type='taxa', color="Genus", point_size = 1, alpha= 0.5, line_alpha = 0.4, label= NULL, title = expression(paste("Ambient ", CO[2])))

p.16s.amb= p.16s.amb + theme(legend.position = "none")


#make legend edits and save to additional plot
bac.leg= p.16s.amb + 
 labs(color = "") +
  theme_void(base_size = 10) + guides(color = guide_legend(override.aes = list(size=3)))
```


```{r fig.width=15, fig.height=10}
# Extract the legend. Returns a gtable
bac.leg.extract <- ggpubr::get_legend(bac.leg)

# Convert to a ggplot and print
bac.leg.plot=as_ggplot(bac.leg.extract) 
```

Make a multiplot
```{r}
bac1= plot_grid(p.16s.amb, p.16s.el, labels = c('A', 'B'), hjust = 0.5)

bac1.final= plot_grid(bac1, NULL, bac.leg.plot, ncol=1, align="v", rel_heights = c(1, 0, 1))

ggsave(paste0(output.loc,"/figure3.jpg"), bac1.final, height= 10)
```


## Compare networks

Start with nets already made from `spieceasi`

```{r}
# Make association matrix
# its1
assoMat.its1.el <- as.matrix(SpiecEasi::symBeta(getOptBeta(se.elevated.its1)))
assoMat.its1.am <- as.matrix(SpiecEasi::symBeta(getOptBeta(se.amb.its1)))

rownames(assoMat.its1.el) <- colnames(assoMat.its1.el) <- colnames(se.elevated.its1$est$data)
rownames(assoMat.its1.am) <- colnames(assoMat.its1.am) <- colnames(se.amb.its1$est$data)


net_its1_comp <- net_asso <- netConstruct(data = assoMat.its1.el, data2= assoMat.its1.am,
                         dataType = "condDependence",
                         sparsMethod = "none",
                         verbose = 0)

# its2
assoMat.its2.el <- as.matrix(SpiecEasi::symBeta(getOptBeta(se.elevated.its2.bstars)))
assoMat.its2.am <- as.matrix(SpiecEasi::symBeta(getOptBeta(se.amb.its2.bstars)))

rownames(assoMat.its2.el) <- colnames(assoMat.its2.el) <- colnames(se.elevated.its2.bstars$est$data)
rownames(assoMat.its2.am) <- colnames(assoMat.its2.am) <- colnames(se.amb.its2.bstars$est$data)


net_its2_comp <- net_asso <- netConstruct(data = assoMat.its2.el, data2= assoMat.its2.am,
                         dataType = "condDependence",
                         sparsMethod = "none",
                         verbose = 0)
# 16s
assoMat.bac.el <- as.matrix(SpiecEasi::symBeta(getOptBeta(se.elevated.bac)))
assoMat.bac.am <- as.matrix(SpiecEasi::symBeta(getOptBeta(se.amb.bac)))

rownames(assoMat.bac.el) <- colnames(assoMat.bac.el) <- colnames(se.elevated.bac$est$data)
rownames(assoMat.bac.am) <- colnames(assoMat.bac.am) <- colnames(se.amb.bac$est$data)


net_16s_comp <- net_asso <- netConstruct(data = assoMat.bac.el, data2= assoMat.bac.am,
                         dataType = "condDependence",
                         sparsMethod = "none",
                         verbose = 0)
```

## Analyze network properties

Determine network properties for objects of class microNet
```{r}
# Using eigenvector centrality as hub score

#its1
comp_its1_props1 <- netAnalyze(net_its1_comp, clustMethod = "cluster_fast_greedy", hubPar = "eigenvector")

plot(comp_its1_props1, groupNames = c("Elevated", "Ambient"))

outputits1.comp.props= summary(comp_its1_props1, groupNames = c("Elevated", "Ambient"))

#its2

comp_its2_props1 <- netAnalyze(net_its2_comp, clustMethod = "cluster_fast_greedy", hubPar = "eigenvector")

plot(comp_its2_props1, groupNames = c("Elevated", "Ambient"))

outputits2.comp.props= summary(comp_its2_props1, groupNames = c("Elevated", "Ambient"))

## hub pars: "degree", "betweenness", "closeness"
# comp_its2_props1.hubpars <- netAnalyze(net_its2_comp, clustMethod = "cluster_fast_greedy", hubPar = c("degree", "betweenness", "closeness"))
# 
# 
# outputits2.comp.props.hubpars= summary(comp_its2_props1.hubpars, groupNames = c("Elevated", "Ambient"))
# 
# hubs.pars.its2=outputits2.comp.props.hubpars$hubs


#16s
comp_16s_props1 <- netAnalyze(net_16s_comp, clustMethod = "cluster_fast_greedy", hubPar = "eigenvector")

plot(comp_16s_props1, groupNames = c("Elevated", "Ambient"))

output16s.comp.props= summary(comp_16s_props1, groupNames = c("Elevated", "Ambient"))
```


## Quantitative network comparison
```{r}
#its1
comp_its1 <- netCompare(comp_its1_props1, 
                          permTest = FALSE, 
                          verbose = FALSE,
                          seed = 123456)

sum.comp.its1=summary(comp_its1, 
        groupNames = c("Elevated CO2", "Ambient CO2"),
        showCentr = c("degree", "between", "closeness"), 
        numbNodes = 5)

#its2
comp_its2 <- netCompare(comp_its2_props1, 
                          permTest = FALSE, 
                          verbose = FALSE,
                          seed = 123456)

sum.comp.its2=summary(comp_its2, 
        groupNames = c("Elevated CO2", "Ambient CO2"),
        showCentr = c("degree", "between", "closeness"), 
        numbNodes = 5)

sum.comp.its2$jaccmat

#16s
comp_bac <- netCompare(comp_16s_props1, 
                          permTest = FALSE, 
                          verbose = FALSE,
                          seed = 123456)

sum.comp.bac=summary(comp_bac, 
        groupNames = c("Elevated CO2", "Ambient CO2"),
        showCentr = c("degree", "between", "closeness"), 
        numbNodes = 5)

sum.comp.bac$jaccmat
```

## Make / output tables of relevant metrics
```{r}
#hubs
hubs.bac= output16s.comp.props$hubs
hubs.bac$dataset <- "16s"

hubs.its1=outputits1.comp.props$hubs
hubs.its1$dataset <- "ITS1"

hubs.its2=outputits2.comp.props$hubs
hubs.its2$dataset <- "ITS2"

hubs= rbind(hubs.bac, hubs.its1)

hubs= rbind(hubs, hubs.its2)

write.csv(hubs, paste0(output.loc, "/hub.taxa.all.csv"))

# netcomp
netmets= sum.comp.its1$propdiffs[1:5,]
netmets$dataset <- "ITS1"
netmets=netmets %>% rownames_to_column("metric")

netmets.its2= sum.comp.its2$propdiffs[1:5,]
netmets.its2$dataset <- "ITS2"
netmets.its2=netmets.its2 %>% rownames_to_column("metric")

netmets.bac=sum.comp.bac$propdiffs[1:5,]
netmets.bac$dataset <- "16s"
netmets.bac= netmets.bac %>% rownames_to_column("metric")

netmets= rbind(netmets, netmets.bac)

netmets= rbind(netmets, netmets.its2)


colnames(netmets) <- c("metric", "Elevated CO2" , "Ambient CO2" ,"difference", "dataset")

write.csv(netmets, paste0(output.loc, "/net.metrics.all.csv"))

#jaccard
jac.mat= sum.comp.its2$jaccmat
jac.mat$dataset <-"its2"
jac.mat= jac.mat %>% rownames_to_column("metric")

jac.its1=sum.comp.its1$jaccmat
jac.its1$dataset <-"its1"
jac.its1= jac.its1 %>% rownames_to_column("metric")

jac.bac=sum.comp.bac$jaccmat
jac.bac$dataset <-"bac"
jac.bac= jac.bac %>% rownames_to_column("metric")

jac.mat= rbind(jac.its1,jac.mat)

jac.mat= rbind(jac.mat, jac.bac)

write.csv(jac.mat, paste0(output.loc, "/jaccard.mat.all.csv"))

```


# SUPPLEMENTAL

Compare ITS primers at the Genus and Family levels

```{r}
its1.gen.unique=  unique(its1.tax$Genus)
its2.gen.unique=  unique(its2.tax$Genus)


its1.fam.unique=  unique(its1.tax$Family)
its2.fam.unique=  unique(its2.tax$Family)

#remove NA
its1.gen.unique= its1.gen.unique[!is.na(its1.gen.unique)]
its2.gen.unique= its2.gen.unique[!is.na(its2.gen.unique)]

its1.fam.unique= its1.fam.unique[!is.na(its1.fam.unique)]
its2.fam.unique= its2.fam.unique[!is.na(its2.fam.unique)]


# look at what is not shared
setdiff(its1.gen.unique, its2.gen.unique) #32 genera found in its1 not in its2
setdiff(its2.gen.unique, its1.gen.unique) #33 general found in its2 not in its1


setdiff(its1.fam.unique, its2.fam.unique) #15 families found in its1 not in its2
setdiff(its2.fam.unique, its1.fam.unique) #20 families found in its2 not in its1


# overlap
overlap.genus=intersect(its1.gen.unique,its2.gen.unique)
overlap.fam=intersect(its1.fam.unique,its2.fam.unique)

#proportion of taxa overlap
length(overlap.genus)/(length(overlap.genus) + length(setdiff(its1.gen.unique, its2.gen.unique))+ length(
setdiff(its2.gen.unique, its1.gen.unique))) # 47.81% taxa are shared at the genus level

length(overlap.fam)/(length(overlap.fam) + length(setdiff(its1.fam.unique, its2.fam.unique))+ length(
setdiff(its2.fam.unique, its1.fam.unique))) # 55.70% taxa are shared at the fam level
```

Venn diagrams

```{r, fig.width= 8}

fam.list=list(ITS1= its1.fam.unique, ITS2= its2.fam.unique)
genera.list=list(ITS1= its1.gen.unique, ITS2= its2.gen.unique)

fam.venn=ggVennDiagram(fam.list) + 
  scale_fill_gradient(low="grey90",high = "red") +
  labs(title = "Taxa overlap by family")

genera.venn=ggVennDiagram(genera.list,) + scale_fill_gradient(low="grey90",high = "red") +
  labs(title = "Taxa overlap by genera")

ggsave(paste0(output.loc,"/figureS1.fam.venn.pdf"), fam.venn)
ggsave(paste0(output.loc,"/figureS2.genera.venn.pdf"), genera.venn)

```
